Trajectory Mapping Results

### 2021-02-02 scripts/analyze_trajectories_v2.Rmd
eu10 and use jar for beast apps

```{r debugging}
# Debugging
args$burnin = 0.1
args$demes = read.csv("./files/demes.csv", stringsAsFactors = FALSE)
args$input = "./results/europeC/201106_europe5.TL.traj"
args$metadata =  "./results/europeC/final_metadata.tsv"
args$n = 50
demes = args$demes
demes$exclude_country[5] = "France,Germany,Italy,Spain"
demes$deme[3] = "China"
demes$division[3] = NA
demes <- demes %>% arrange(deme)
```

```{r plotopts}
# Plot options -----------------------------------------------------------------------
set_plotopts()
dcolors <- pal_npg("nrc")(10)[c(1:5,9)]
dcolors <- dcolors[1:nrow(demes)]
names(dcolors) <- sort(demes$deme)
dark_dcolors <- rgb(t(col2rgb(dcolors)/2), maxColorValue=255)
names(dark_dcolors) <- names(dcolors)
ecolors <- c("#1e3d59", "#f5f0e1", "#ff6e40", "#ffc13b")
names(ecolors) <- c("B", "M", "IM", "OM")
```

@@

zoom_to_xy <- st_transform(st_sfc(st_point(zoom_to), crs = 4326),
                           crs = target_crs)
disp_window <- st_sfc(
    st_point(st_coordinates(zoom_to_xy - c(x_span/2, y_span/2))),
    st_point(st_coordinates(zoom_to_xy + c(x_span *1.2, y_span *1.2))),
    crs = target_crs
)
ggplot() + geom_sf(data = worldmap, aes(fill = deme), alpha = 0.8, size = 0) +
    scale_x_continuous(breaks = seq(-100, 200, by = 10)) +
    scale_y_continuous(breaks = seq(0, 100, by = 10)) +
    #geom_sf(data = zoom_to_xy, color = 'red') +
    coord_sf(xlim = st_coordinates(disp_window)[,'X'],
             ylim = st_coordinates(disp_window)[,'Y'],
             crs = target_crs, expand = FALSE) +
  scale_fill_manual(values = dcolors, na.value = "grey90") +
  theme_void() +
  theme(panel.grid.major = element_line(colour = "grey", size = 0.1),
        legend.position = "none")
        # panel.ontop = TRUE,
        # panel.background = element_blank())

@@

case_data <- get_cases(demes, from = ymd("2019-09-01"), to = MRS + 10) %>%
  arrange(date) %>%
  group_by(deme) %>%
  rename(value = cases) %>%
  mutate(cumvalue = cumsum(value)) %>%
  group_by(deme, date) %>%
  mutate(cumvalue = max(cumvalue)) %>%
  select(deme, date, cumvalue, value)  %>%
  distinct_at(vars(deme, date, cumvalue), .keep_all = TRUE)

@@

case_data_pop <- case_data %>% left_join(population, by="deme") %>% mutate(ECDCcas100 = cumcases * 1e5 / (pop * 1e3)) 
```

@@

:::: {.column width="40%"}
\small
```{r ecdc-table, echo=FALSE}
knitr::kable(case_data_pop %>% ungroup %>% filter(date == MRS + 10) %>% select(deme, cumcases, pop, cas100) %>% mutate_if(is.numeric, function (x) round(x,1)), booktabs = TRUE, caption = "Total number of cases reported to ECDC 18th March")
```
::::

:::: {.column width="60%"}
```{r ecdc-plot, fig.height = 3, fig.width=4, fig.cap="ECDC case counts for each deme from the beginning of the pandemic to March 18"}
ggplot(case_data) +
  geom_line(aes(x = date, y = cumcases, color = deme)) +
   scale_color_manual(values = dcolors, name = "") +
   scale_x_date(limits = c(ymd("2019-12-30"), ymd("2020-03-18")), date_breaks = "1 month", date_labels = "%b %d") +
    scale_y_log10(labels = trans_format("log10", math_format(10^.x))) +
    ylab("Case count") +
    theme(axis.title.x=element_blank(),
          legend.position = c(0.2, 0.7),
          legend.box="horizontal",
          panel.grid.major.x =  element_line(colour = "grey80", size = 0.3),
          panel.grid.minor.x =  element_line(colour = "grey80", size = 0.2))
```

@@

# Do this in a function, useful also for Claire
events <- df$events %>%
  mutate(src = factor(src,
                      levels = 0:(nrow(demes) - 1),
                      #labels = sort(demes$deme)),
                      labels = c("France", "Germany","China","Italy","OtherEuropean","Spain")),
         dest = (factor(dest,
                           levels = 0:(nrow(demes) - 1),
                           #labels = sort(demes$deme),
                           labels = c("France" , "Germany","China","Italy","OtherEuropean","Spain"))))
events[events$event == "O", "src"] = "China"
events[events$event == "O", "mult"] =  1

@@

df_traj_cases <- df_traj %>%
  mutate(traj = as.character(traj)) %>%
  bind_rows(case_data %>%
            filter(cumvalue != 0) %>%
            mutate(var = "ECDC"))
                  # traj = "ecdc"))
```

@@

@@

```{r}
# Active cases
d = "France"
df_traj_cases %>%
    # 1. Prepare data
    filter(deme == d, var %in% c("active_pop", "in_pop", "ECDC"), is.na(partner)) %>%
    # 2. Plot
    ggplot() +
    geom_line(aes(date, cumvalue, group = traj,
                  color = var)) +
                  # color = traj == "ecdc",
                  # alpha = traj == "ecdc",
                  # linetype = traj == "ecdc",
                  # size = traj == "ecdc")) +
    # scale_color_manual(name = "",
    #                    labels = c(paste0(deme, " inferred\n population trajectories"),
    #                               paste0(ifelse(deme == china_deme, "China", as.character(deme)), 
    #                                      " ECDC total case count")),
    #                    values = c(dcolors[[deme]], dark_dcolors[[deme]])) +
    # scale_alpha_manual(name = "", values = c(0.2, 1), guide = FALSE) +
    # scale_linetype_manual(name = "", values = c(1, 2),
    #                       labels = c(paste0(deme, " inferred\n population trajectories"),
    #                                  paste0(ifelse(deme == china_deme, "China", as.character(deme)), 
    #                                         " ECDC total case count"))) +
    # scale_size_manual(name = "", values = c(0.5, 1),
    #                   labels = c(paste0(deme, " inferred\n population trajectories"),
    #                              paste0(ifelse(deme == china_deme, "China", as.character(deme)), 
    #                                     " ECDC total case count"))) +
    scale_x_date(limits = c(ymd("2019-11-01"), ymd(MRS + 10)), date_breaks = "2 weeks", 
                 date_labels = "%b %d") +
    scale_y_log10(expand = c(0, 0),labels = trans_format("log10", math_format(10^.x))) +
    ylab("Population size") +
    theme(legend.position = c(0.4, 0.9),
          axis.title.x=element_blank())
```

>>>>

```{r debugging}
# Debugging
args$burnin = 0.1
args$demes = read.csv("./files/demes_China.csv", stringsAsFactors = FALSE)
#args$input = "./results/europeC/201211_europe6p2.TL.traj"
args$input = "/Users/maceci/traj_10_test/notimeparameter/210122_europe10a.TL.traj"
args$metadata =  "./results/europeC/final_metadata.tsv"
args$n = 200
demes = args$demes
demes$exclude_country[5] = "France,Germany,Italy,Spain"
#demes$deme[3] = "China"
#demes$division[3] = NA
demes <- demes %>% arrange(deme)
```

```{r plotopts}
# Plot options -----------------------------------------------------------------------
set_plotopts()
dcolors <- pal_npg("nrc")(10)[c(1:5,9)]
dcolors <- c("#E64B35FF", "#4DBBD5FF", "#00A087FF", "#3C5488FF", "#ffaa4fFF", "#7E6148FF")
#dcolors <- c("#E64B35FF", "#7E6148FF", "#00A087FF", "#3C5488FF", "#4DBBD5FF", "#ffc954")
#dcolors <- dcolors[1:nrow(demes)]
names(dcolors) <- sort(demes$deme)
dark_dcolors <- rgb(t(col2rgb(dcolors)/2), maxColorValue=255)
names(dark_dcolors) <- names(dcolors)
ecolors <- c(B = "#1e3d59", M = "#f5f0e1", IM = "#ff6e40", OM = "#ffc13b")
```

@@

zoom_to_xy <- st_transform(st_sfc(st_point(zoom_to), crs = 4326),
                           crs = target_crs)
disp_window <- st_sfc(
    st_point(st_coordinates(zoom_to_xy - c(x_span/2, y_span/2))),
    st_point(st_coordinates(zoom_to_xy + c(x_span *1.2, y_span *1.25))),
    crs = target_crs
)
ggplot() + geom_sf(data = worldmap, aes(fill = deme), alpha = 0.9, size = 0) +
    scale_x_continuous(breaks = seq(-100, 200, by = 10)) +
    scale_y_continuous(breaks = seq(0, 100, by = 10)) +
    #geom_sf(data = zoom_to_xy, color = 'red') +
    coord_sf(xlim = st_coordinates(disp_window)[,'X'],
             ylim = st_coordinates(disp_window)[,'Y'],
             crs = target_crs, expand = FALSE) +
  scale_fill_manual(values = dcolors, na.value = "grey90") +
  #scale_fill_manual(values = c("#E64B35FF", "grey50", "grey50", "grey50", "grey50", "grey50"), na.value = "grey90") +
  #ggsci::scale_fill_tron()
  theme_void() +
  theme(panel.grid.major = element_line(colour = "grey", size = 0.1),
        legend.position = "none")
        # panel.ontop = TRUE,
        # panel.background = element_blank())

@@

case_data <- get_cases(demes, from = ymd("2019-09-01"), to = MRS + 10) %>%
  arrange(date) %>%
  group_by(deme, date) %>%
  rename(value = new_cases) %>%
  summarise(value = sum(value)) %>%
  mutate(cumvalue = cumsum(value))

library(zoo)

case_data <- case_data %>%
  group_by(deme) %>%
  mutate(rollavg3 = zoo::rollmean(value, k = 3, fill = NA),
         #rollavg7 = zoo::rollmean(value, k = 7, fill = NA),
         datereport = date,
         date = date - 10)

@@

case_data_pop <- case_data %>% left_join(population, by="deme") %>% mutate(ECDCcas100 = cumvalue * 1e5 / (pop * 1e3)) 
```

@@

:::: {.column width="40%"}
\small
```{r ecdc-table, echo=FALSE}
knitr::kable(case_data_pop %>% ungroup %>% filter(date == MRS + 10) %>% select(deme, cumvalue, pop, ECDCcas100) %>% mutate_if(is.numeric, function (x) round(x,1)), booktabs = TRUE, caption = "Total number of cases reported to ECDC 18th March")
```
::::

:::: {.column width="60%"}
```{r ecdc-plot, fig.height = 3, fig.width=4, fig.cap="ECDC case counts for each deme from the beginning of the pandemic to March 18"}

ggplot(case_data) +
  geom_line(aes(x = date, y = cumvalue, color = deme)) +
  geom_point(aes(x = date, y = cumvalue, color = deme)) +
   scale_color_manual(values = dcolors, name = "") +
   scale_x_date(limits = c(ymd("2019-12-30"), ymd("2020-03-18")), date_breaks = "1 month", date_labels = "%b %d") +
  scale_y_log10(labels = trans_format("log10", math_format(10^.x))) +
  #scale_y_continuous(expand = c(0, 0)) +
    ylab("Case count") +
    theme(axis.title.x=element_blank(),
          legend.position = c(0.2, 0.7),
          legend.box="horizontal",
          panel.grid.major.y =  element_line(colour = "grey80", size = 0.3, linetype = 2),
          panel.grid.minor.y =  element_line(colour = "grey80", size = 0.2,  linetype = 2),
          panel.background =  element_blank())
```

@@

# Do this in a function, useful also for Claire
events <- df$events %>%
  mutate(src = factor(src,
                      levels = 0:(nrow(demes) - 1),
                      labels = sort(demes$deme)),
                      #labels = c("France", "Germany","China","Italy","OtherEuropean","Spain")),
         dest = (factor(dest,
                           levels = 0:(nrow(demes) - 1),
                           labels = sort(demes$deme))))
                           #labels = c("France" , "Germany","China","Italy","OtherEuropean","Spain"))))
events[events$event == "O", "src"] = "China"
events[events$event == "O", "mult"] =  1

@@

df_traj_cases <- df_traj %>%
  mutate(traj = as.character(traj)) %>%
  bind_rows(case_data %>%
            #filter(cumvalue != 0) %>%
            mutate(var = "ECDC"))
                  # traj = "ecdc"))

min_date <- min(df_traj_cases$date)
df_traj_cases_complete <- df_traj_cases %>%
  group_by(traj, var, deme, partner) %>%
  complete(date = seq.Date(min_date, MRS, by = "day")) %>%
  replace_na(list(value = 0)) %>%
  arrange(traj, date, deme) %>%
  mutate(cumvalue = cumsum(value))
```

@@

# r_reported <- reported %>%
#   mutate(Country = type,
#          `Cases ECDC` = cumcases,
#          `Median Cases (95% CCI)` = paste0(round(Imedian), " (", round(Ilow), "-", round(Ihigh), ")"),
#          `Median Reporting Rate (95% CCI)` = paste0(round(reportrate, 2), " (", round(hreportrate, 2), "-", round(lreportrate, 2), ")")) %>%
#   select(-one_of(colnames(reported)))
#
# write_tsv(r_reported, paste0(args$output_table, "01.tsv"))

reported1 <- df_traj_cases %>%
  filter(var == "in_pop", date == MRS) %>%
  left_join(case_data_pop %>% rename(cumcases = cumvalue) %>% select(deme, date, cumcases, pop), by = c("deme", "date")) %>%
  group_by(var, deme, cumcases, pop) %>%
  summarise(median_in = median(cumvalue),
            l95_in = HDInterval::hdi(cumvalue, 0.95)[1],
            h95_in = HDInterval::hdi(cumvalue, 0.95)[2]) %>%
  mutate(times_med = median_in/cumcases,
         times_l = l95_in/cumcases,
         times_h = h95_in/cumcases,
         p_med = median_in/(pop*1e3) * 100,
         p_l = l95_in/(pop*1e3) * 100,
         p_h = h95_in/(pop*1e3) * 100)

reported2 <- df_traj_cases %>%
  filter(var == "in_pop") %>%
  left_join(case_data_pop %>% rename(cases = value, ecdcavg3=rollavg3) %>% select(deme, date, cases, pop, ecdcavg3), by = c("deme", "date")) %>%
  group_by(var, deme, date, cases, ecdcavg3) %>%
  summarise(median_in = median(value),
            l95_in = HDInterval::hdi(value, 0.95)[1],
            h95_in = HDInterval::hdi(value, 0.95)[2]) %>%
  mutate(report_rate = cases / median_in)

reported3 <- df_traj_cases %>%
  filter(var == "in_pop") %>%
  left_join(case_data_pop %>% rename(cases = value, ecdcavg3=rollavg3) %>% select(deme, date, cases, pop, ecdcavg3), by = c("deme", "date")) %>%
  group_by(var, deme, week = week(date), cases, ecdcavg3) %>%
  summarise(median_in = median(value),
            l95_in = HDInterval::hdi(value, 0.95)[1],
            h95_in = HDInterval::hdi(value, 0.95)[2]) %>%
  mutate(report_rate = cases / median_in)

reported3 <- df_traj_cases_complete %>%
  group_by(var, deme, week = week(date), cases, ecdcavg3) %>%
  summarise(median_in = median(value),
            l95_in = HDInterval::hdi(value, 0.95)[1],
            h95_in = HDInterval::hdi(value, 0.95)[2]) %>%
  mutate(report_rate = cases / median_in)

ggplot(reported3 %>% filter(deme != "China")) +
  geom_line(aes(week, report_rate, color = deme)) +
  geom_point(aes(week, report_rate, color = deme))

@@

```{r}
# Active cases
library(gganimate)
d = "Germany"
df_traj_cases %>%
    # 1. Prepare data
    filter(deme == d, 
           #traj %in% c(1, NA), 
           var %in% c("in_pop", "ECDC"), 
           is.na(partner)) %>%
    # 2. Plot
    ggplot() +
    geom_line(aes(date, cumvalue, group = traj,
                  color = var == "ECDC",
                  alpha = var == "ECDC",
                  linetype = var == "ECDC",
                  size = var == "ECDC")) +
    scale_color_manual(name = "",
                       labels = c(paste0(d, " inferred\n population trajectories"),
                                  paste0(ifelse(d == china_deme, "China", as.character(d)),
                                         " ECDC total case count")),
                       values = c(dcolors[[d]], dark_dcolors[[d]])) +
                       #values = c("#4ab9ed", dark_dcolors[[d]])) +
    scale_alpha_manual(name = "", values = c(0.2, 1), guide = FALSE) +
    scale_linetype_manual(name = "", values = c(1, 2),
                          labels = c(paste0(d, " inferred\n population trajectories"),
                                     paste0(ifelse(d == china_deme, "China", as.character(d)),
                                            " ECDC total case count"))) +
    scale_size_manual(name = "", values = c(0.5, 1),
                      labels = c(paste0(d, " inferred\n population trajectories"),
                                 paste0(ifelse(d == china_deme, "China", as.character(d)),
                                        " ECDC total case count"))) +
    scale_x_date(limits = c(ymd("2019-11-01"), ymd(MRS + 10)), date_breaks = "2 weeks",
                 date_labels = "%b %d") +
    scale_y_log10(expand = c(0, 0),labels = trans_format("log10", math_format(10^.x))) +
    ylab("Population size") +
    theme(legend.position = c(0.4, 0.9),
          axis.title.x=element_blank())
  #theme_void() #+
  #transition_reveal(date)
```
